### 2012-09-28 doc/prelim/document/chapters/research_proposal/research_proposal.tex
finished research proposal draft

\section{Challenge Problem}
\label{sec:challenge_problem}
Our numerical experiments will provide us with a set of metrics giving
best-performance solver parameters for problems of different types. As
an additional bonus, these experiments will let us further develop new
parallel implementations of the stochastic methods in order to improve
their performance and scalability. With these advancements in hand, we
look forward towards applying them to large-scale problems in reactor
physics.

\section{Conclusion}
\label{sec:conclusion}

>>>>

\section{Proposed Challenge Problem}
\label{sec:challenge_problem}
Our numerical experiments will provide us with a set of metrics giving
best-performance stochastic solver parameters for problems of
different types. As an additional bonus, these experiments will let us
further develop new parallel implementations of the stochastic methods
in order to improve their performance and scalability. With these
advancements in hand, we look forward towards applying them to
large-scale problems in reactor physics. The Consortium for Advanced
Simulation of Light Water Reactors (CASL) is a Department of Energy
Modeling and Simulation Hub focused on predictive multiphysics
modeling of nuclear reactors
\citep{u.s._depertment_of_energy_casl:_2011}. Their effors are focused
on enabling nuclear reactor analysis and development using today's
supercomputing technology to achieve higher power uprates, higher fuel
burnup, and predictive accident scenario analysis. As a means of
driving their research and development, CASL has worked with the
nuclear industry to identify challenge problems of critical importance
to achieving these goals.

To meet some of the high fidelity modeling requirements of these
challenge problems, the Drekar multiphysics code is being developed at
Sandia National Laboratories for fluid flow and conjugate heat
transfer problems \citep{pawlowski_drekar::cfd_2012}. Drekar is being
used to analyze complex problems including the grid-to-rod-fretting
(GTRF) phenomena observed in light water reactor fuel rods. GTRF
occurs due to the turbulent flow of water through the subchannels of
the reactor core. This turbulence causes the nuclear fuel pins to
vibrate and the fuel cladding to wear against structures in the
reactor, causing localized hot spots and ultimately cladding
failure. Reducing these instances of failure and understanding how
they occur through predictive multiple physics simulation of fluid
flow and heat transfer will permit more robust fuel assembly
designs. In addition, the turbulence studies that Drekar performs will
allow for improved flow patterns to be generated in the assembly
subchannels, allowing for better heat transfer between the nuclear
fuel and the water, ultimately allowing for a more efficient reactor.

Drekar is a massively parallel physics framework built upon the same
Panzer library as well as other open-source tools that we are
leveraging in our own experimental framework. Furthermore, its
multiphysics formulation utilizes Newton-Krylov methods to generate a
fully consistent solution scheme for the Navier-Stokes equations with
various turbulence models for fluid simulation as well as the heat
conduction problem for simulation in solid materials. Performance
metrics show scaling to $O(100,000)$ cores on leadership class
machines with $O(\sn{1}{9})$ degrees of freedom in these
simulations. Given these features, Drekar will be an excellent
framework in which we can pose our own challenge problem for the FANM
and newly parallelized MCSA methods proposed by this research. At the
time our preliminary numerical experiments are completed for
characterizing the parallel FANM and MCSA implementations, we propose
implementing them within the Drekar framework and performing the
equivalent of its largest GTRF mutliphysics problem to date, most
likely to be a fuel-assembly sized problem. Doing so will allow us to
not only implement our new methods in a production physics code and
apply it to a problem of great interest to the nuclear community, but
it will also allow us to compare its performance and scaling to the
Newton-Krylov methods currently leveraged within Drekar. If
successful, we expect to demonstrate improved computational
performance from a memory usage and/or an algorithmic scaling
perspective.

\section{Conclusion}
\label{sec:conclusion}
We have proposed the research and development of novel, massively
parallel stochastic methods for the solution of linear and nonlinear
problems. These new methods aim to enhance the solution of multiple
physics problems of great national interest. Current progress has
shown that these methods can indeed be expanded into a general linear
algebra framework that permits their application to a broad range of
problems. Additional analysis has also been completed that begins to
characterize their performance in different configurations. Leveraging
this work, we have defined a multiphysics framework based on the
finite element method and have demonstrated that these stochastic
methods can be used effectively to solve the linear systems generated
by these discretizations.

To guide our development of parallel stochastic methods, we have
devised several classes of numerical experiments using our
multiphysics framework that will provide insight into their properties
and performance for linear and nonlinear problems. Furthermore, these
experiments aim to provide metrics that will give an indication of
both solver performance for various problem types and the type of
parallel decomposition required to effectively solve these
problems. Domain decomposition, domain replication, and their
combination will be of great interest for the analysis of parallel
Monte Carlo methods while algorithm scalability and memory usage will
be of interest for the MCSA and FANM methods that leverage them.

We culminate our research and development with a challenge problem by
applying these new stochastic methods in the Drekar multiphysics
code. With Drekar, we look forward to providing a new and robust
solution method at the production scale. With these tools and a
leadership class computing platform, we will provide performance
measurements for the stochastic solvers and compare them to
conventional methods currently leveraged within Drekar. With the
potential to improve computational performance in these environments,
these new methods have the potential to help meet the high fidelty
modeling and simulation needs of the nuclear community.
